---
title: "covid 19"
author: "Marcia Alarcon"
date: "10/12/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: Dying from Covid-19 is independent of sex.
$H_A$: Dying from Covid-19 is dependent of sex.



Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Among those who died in Korea from Covid-19 were mostly men. Women died at almost half the rate as men. The right bar graph shows that men(blue) died at about 63% of those infected and the women(red) died at about a 36% of those infected.


  





 
  
## Numerical results

```{r}
table2 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table2)
colPerc(table2)
```

Men died at almost twice the rate as woman did from Covid-19 in Korea. In the first chart 36% of the dead count were females and 64% of the death count was men.

Very few people actually died of Covid-19, 1.4% of Korean females died however 3% of Korean men died. Therefore, Korean men die almost twice as more then the women rate.

## Inferential results 

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The p-value 0.001 in the Pearson's chi-squared test is small enough to reject the Null hypothesis. The death rate in Korean from Covid-19 is dependent on sex.

The Fisher Exact test shows that men died at 2.3 times than the rate women die.

## Conclusions

The question was if dying is independent of sex for Covid-19. After analyzing the ggplot, numerical summary, Pearson's chi-square and the Fisher Exact test we reject the Null Hypothesis because the p-value is so small. The results show that Men died at 2.3 times the rate that women died, resulting that the death rate of Korean were dependent on sex for Covid-19.